Comparing human and model-based forecasts of COVID-19 in Germany and Poland

### 2021-10-14 manuscript/manuscript.Rmd
update abstract

Model-based forecasts, which have played an important role in shaping public policy throughout the Covid-19 pandemic, represent an implicit combination of model assumptions and the researcher’s subjective opinion. This work explicitly separates human opinion and model-derived insights to assess and compare the relative strengths and weaknesses of both approaches. We compared purely opinion-derived forecasts of cases and deaths from COVID-19 in Germany and Poland, elicited from researchers and volunteers (“crowd forecasts”), against predictions from two semi-mechanistic epidemiological models. In addition, we compared these forecasts against an ensemble of model-based, but expert-tuned, forecasts, submitted to the German and Polish Forecast Hub by other research institutions (“Hub ensemble”). We found an ensemble of crowd-sourced predictions to outperform all other methods when predicting cases (Weighted Interval Score relative to the Hub ensemble at a two-week horizon: 0.89), but not when predicting deaths (rel. WIS 1.26). Crowd forecasts were severely overconfident (55% and 75% coverage of the 90% prediction intervals for cases and deaths, respectively) when compared to model-based predictions. Performance of the untuned models deteriorated quickly with increasing forecast horizon when assumptions were no longer met (rel. WIS > 1.64 for all four week ahead predictions).

@@

Data on daily reported test positive cases and deaths linked to Covid-19 were provided by the organisers of the German and Polish forecast hub. Until December 14th 2020 these data were sourced from the European Centre for Disease Control (ECDC) [@DownloadHistoricalData2020a]. After ECDC stopped publishing daily data, observations were sourced from the Robert Koch Institute (RKI) and the Polish Ministry of Health for the remainder of the submission period  [@RKICoronavirusSARSCoV2a]. These data are subject to reporting artefacts (such as a delayed case reporting in Poland on the 24th November [@RozbieznosciStatystykachKoronawirusa0100]), changes in reporting over time and variation in testing regimes (e.g. in Germany from the 11th of November on [@aerzteblattSARSCoV2DiagnostikRKIPasst2020]).

>>>>

Model-based forecasts, which have played an important role in shaping public policy throughout the COVID-19 pandemic, represent an implicit combination of model assumptions and the researcher’s subjective opinion. This work analyses and compares human opinion against model-derived insights to discern relative strengths and weaknesses of both approaches. We compared purely opinion-derived forecasts of cases and deaths from COVID-19 in Germany and Poland, elicited from researchers and volunteers, against predictions from two semi-mechanistic epidemiological models. We also compared these forecasts against an ensemble of model-based, but expert-tuned, forecasts, submitted to the German and Polish Forecast Hub by other research institutions. In addition, we examined the effects of our contributions to the performance of the Hub ensemble. We found aggregated crowd forecasts to outperform all other methods when predicting cases (Weighted Interval Score relative to the Hub ensemble: 0.89), but not when predicting deaths (rel. WIS 1.26). Crowd forecasts were noticeably more overconfident (55% and 75% coverage of the 90% prediction intervals for cases and deaths, respectively) than model-based predictions. Performance of the semi-mechanistic models was good short-term, but deteriorated quickly over time when assumptions were no longer met.

@@

Data on daily reported test positive cases and deaths linked to COVID-19 were provided by the organisers of the German and Polish forecast hub. Until December 14th 2020 these data were sourced from the European Centre for Disease Control (ECDC) [@DownloadHistoricalData2020a]. After ECDC stopped publishing daily data, observations were sourced from the Robert Koch Institute (RKI) and the Polish Ministry of Health for the remainder of the submission period  [@RKICoronavirusSARSCoV2a]. These data are subject to reporting artefacts (such as a delayed case reporting in Poland on the 24th November [@RozbieznosciStatystykachKoronawirusa0100]), changes in reporting over time and variation in testing regimes (e.g. in Germany from the 11th of November on [@aerzteblattSARSCoV2DiagnostikRKIPasst2020]).
